# R basics

## Using R as a calculator


R supports a  range of mathematical operators, most application in R will involve some numeric calculation. 
```{r}
1+2
```


```{r}
40*4-10
```

## Saving variables 

To save a variable we use <- to assign a certain number a 

Here we assign the radius of a circle to a value of 40.

```{r}
radius <- 40
```


We can use these saved values in later calculations. 
In this example we can calculate the area of a circle with the raduis from before. 

```{r}
area <- radius*pi*2
area
```


::: {#hello .greeting .message width="40%"}
Hello **world**!
:::


!!!Challenge : use the R command line to calculate the area of a rectangle that has side lengths 30 and 20


## Strings
Strings store a collection of characters such as names, places or anything that cant be expressed by a number. Strings are declared by using either single or double quotation marks. 

```{r}
myName <- "Josh"
```

Strings can be stuck together using the paste function
```{r}
greeting <- paste("Hello, my name is",myName)
greeting
```

## Booleans

Booleans store only true or false values. If a boolean is used to keep a true or false value. Booleans are also great for
```{r}
myHeight <- 178
#if taller than 120, then attendee can go on the park ride
canRide <- myHeight > 120
canRide
```


## Vectors

Vectors are used to save a collection of variables, they can be only used to save data of the same type.

Create a vector using the c() function, this will store all varialbles in the same place, allowing us to refrence them as a single unit in R
```{r}
numbers = c(5,6,2,1,7)
```
In order to take a number from a vector we must use indexing to obtain the item from the vector that is in that position
Vectors are indexed with square brackets, starting with one. In


```{r}
numbers[1]
```
```{r}
numbers[2:4]
```
```{r}
numbers[-3]
```


Logical indexing means we can obtain true or false values for each 
the follwing code 
```{r}
numbers>5

```


## Lists

Lists are very similar to vectors, but can store any type of data that is available. In a list we can keep strings, booleans, functions and even enire models and dataframes.

Indexing a list works much like indexing a vector, however we will return a list with that element in it. 
```{r}

```
In order to in
```{r}

```


    try storing some lists in a dataframe 
## Loops*

Optional cover if we have time

```{r}

```

## Conditional Statements* 
```{r}

```



## Functions*

```{r}

```

